---
title: "Vecchia_FIC_PIC_compare_est_parms"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preparation}
rm(list = ls())
set.seed(123)
library(GpGp)
library(mvtnorm)
library(parallel)

n <- 5000
d <- 10
mVec <- seq(from = 4, to = 52, by = 4)
nrep <- 10
rTrue <- 20
rFalse <- 0.2
r <- c(rep(rTrue, 2), rep(rFalse, d - 2))
sigmaSq <- 1
tauSq <- 0
kernel_func <- matern25_scaledim_relevance
kernelName <- "matern25_scaledim_relevance"
```

## Function for simulating replicates
Function for simulating predictors and GP. Return a list consisting of locations `locs` and observations `y`.
```{r func for simulating predictors and GP}
simulate_func <- function(nExtra)
{
  nTtl <- n + nExtra
  locs <- lhs::randomLHS(nTtl, d)
  # locs <- locs * outer(rep(sqrt(nTtl), nTtl), 1 / sqrt(colSums(locs^2)))
  covM <- kernel_func(c(sigmaSq, r, tauSq), locs)
  cholM <- t(chol(covM))
  y <- as.vector(cholM %*% rnorm(nTtl))
  list(locs = locs, y = y)
}
```

## FIC and PIC
Function for selecting the inducing points (pseudo-inputs). Return the indices of inducing points.
```{r func for selecting inducing pts}
select_induce <- function(locs, m)
{
  order_maxmin(locs)[1 : m]
}
```
Function for partitioning the observations based on the `kmeans` function.
```{r func for partitioning}
partition_func <- function(locs, k)
{
  clusterID <- kmeans(locs, centers = k)$cluster
  partition <- list()
  for(i in 1 : k)
    partition[[i]] <- which(clusterID == i)
  partition
}
```
Function for computing the low-rank covariance matrices, returning the sum of the low-rank covariance matrix `Q` and the diagonal nugget.
```{r func for lr covM}
lr_covM <- function(locs, centers, theta)
{
  covMTTl <- kernel_func(theta, rbind(locs, centers))
  m1 <- nrow(locs)
  m2 <- nrow(centers)
  M <- covMTTl[(m1 + 1) : (m1 + m2), (m1 + 1) : (m1 + m2)]
  KXM <- covMTTl[1 : m1, (m1 + 1) : (m1 + m2)]
  Q <- KXM %*% solve(M) %*% t(KXM)
  diag(Q) <- diag(covMTTl)[1 : m1]
  Q
}
```
Function for the FIC log-likelihood. `locs` and `centers` should not overlap.
```{r func FIC log-likelihood}
loglk_FIC <- function(y, locs, centers, theta = c(sigmaSq, r, tauSq))
{
  Q <- lr_covM(locs, centers, theta)
  dmvnorm(y, sigma = Q, log = T)
}
```
Function for the PIC log-likelihood. `locs` and `centers` should not overlap.
```{r func PIC log-likelihood}
loglk_PIC <- function(y, locs, centers, theta = c(sigmaSq, r, tauSq))
{
  partition <- partition_func(locs,
                              round(nrow(locs) / nrow(centers)))
  Q <- lr_covM(locs, centers, theta)
  for(i in 1 : length(partition))
  {
    setI <- partition[[i]]
    if(length(setI) > 0)
    {
      Q[setI, setI] <- kernel_func(theta, 
                                   locs[setI, , drop = F])
    }
  }
  dmvnorm(y, sigma = Q, log = T)
}
```

## Vecchia and Scaled Vecchia
Function for the Vecchia log-likelihood.
```{r func Vecchia log-likelihood}
loglk_Vecchia <- function(y, locs, m, theta = c(sigmaSq, r, tauSq))
{
  odr <- order_maxmin(locs)
  locsOdr <- locs[odr, ]
  NNarray <- find_ordered_nn(locsOdr, m)
  yOdr <- y[odr]
  vecchia_meanzero_loglik(theta, kernelName,
                          yOdr, locsOdr, NNarray)[[1]]
}
```
Function for the scaled Vecchia log-likelihood
```{r func scaled Vecchia log-likelihood}
loglk_SVecchia <- function(y, locs, m, theta = c(sigmaSq, r, tauSq))
{
  r <- theta[2 : (d + 1)]
  odr <- order_maxmin(t(t(locs) * r))
  locsOdr <- locs[odr, ]
  NNarray <- find_ordered_nn(t(t(locsOdr) * r), m)
  yOdr <- y[odr]
  vecchia_meanzero_loglik(theta, kernelName, 
                          yOdr, locsOdr, NNarray)[[1]]
}
```

## Simulate locs and y lists
```{r locs and y lists, eval = F}
for(i in 1 : nrep)
{
  mMax <- max(mVec)
  simObj <- simulate_func(mMax)
  locs <- simObj$locs
  idxInduce <- select_induce(locs, mMax)
  locs <- rbind(locs[idxInduce, ], locs[- idxInduce, ])
  
  write.table(locs, file = paste0("locs_", i, ".csv"), row.names = F, 
            col.names = F, sep = ",")
  write.table(simObj$y, file = paste0("y_", i, ".csv"), row.names = F, 
            col.names = F, sep = ",")
}
```

## Simulation
```{r simulation}
rslt <- matrix(NA, 5 * length(mVec) * nrep, 4)
colnames(rslt) <- c("iter", "m", "method", "llk")
rslt[, 3] <- c(1, 2, 3, 4, 5) # method ID
rslt[, 2] <- kronecker(mVec, rep(1, 5)) # m
rslt[, 1] <- kronecker(c(1 : nrep), rep(1, 5 * length(mVec))) # iter ID
thetaInit <- c(1, rep(1, d), 0.1)
thetaLB <- c(1e-4, rep(1e-8, d), 1e-4)

mcObj <- mclapply(1 : nrep, function(i){
  mMax <- max(mVec)
  locs <- as.matrix(read.table(paste0("locs_", i, ".csv"), 
                               header = F, sep = ","))
  idxInduce <- 1 : mMax
  locsCp <- locs
  locs <- locs[- idxInduce, ]
  # true density
  trueDen <- dmvnorm(rep(0, n),
                     sigma = kernel_func(c(sigmaSq, r, tauSq), locs),
                     log = T)
  retVec <- rep(NA, length(mVec) * 5)
  for(j in 1 : length(mVec)){
    m <- mVec[j]
    locsInduceFIC <- locsCp[idxInduce[1 : m], ]
    locsInducePIC <- locsCp[idxInduce[1 : (m / 2)], ]
    locsInduceFITC <- as.matrix(read.table(paste0("induce_", i, "_", m, ".csv"), 
                               header = F, sep = ","))
    
    loglk_FIC_theta <- function(theta){
      loglk_FIC(rep(0, n), locs, locsInduceFIC, theta)
    }
    loglk_PIC_theta <- function(theta){
      loglk_PIC(rep(0, n), locs, locsInducePIC, theta)
    }
    loglk_Vecchia_theta <- function(theta){
      loglk_Vecchia(rep(0, n), locs, m, theta)
    }
    loglk_SVecchia_theta <- function(theta){
      loglk_SVecchia(rep(0, n), locs, m, theta)
    }
    
    loglk_FIC_opt <- optim(par = thetaInit, fn = loglk_FIC_theta, 
                           method = "L-BFGS-B", lower = thetaLB)$value
    loglk_PIC_opt <- optim(par = thetaInit, fn = loglk_PIC_theta, 
                           method = "L-BFGS-B", lower = thetaLB)$value
    loglk_Vecchia_opt <- optim(par = thetaInit, fn = loglk_Vecchia_theta, 
                           method = "L-BFGS-B", lower = thetaLB)$value
    loglk_SVecchia_opt <- optim(par = thetaInit, fn = loglk_SVecchia_theta, 
                           method = "L-BFGS-B", lower = thetaLB)$value
    
    offset <- (j - 1) * 5
    retVec[offset + 1] <- trueDen - loglk_FIC_opt
    retVec[offset + 2] <- trueDen - loglk_PIC_opt
    retVec[offset + 3] <- trueDen - loglk_Vecchia_opt
    retVec[offset + 4] <- trueDen - loglk_SVecchia_opt
    retVec[offset + 5] <- trueDen - loglk_FIC(rep(0, n), locs, locsInduceFITC)
  }
}, mc.cores = nrep)
for(i in 1 : nrep){
    rslt[((i - 1) * length(mVec) * 5 + 1) : (i * length(mVec) * 5), 4] <- 
      mcObj[[i]]
}
save(rslt, file = "Vecchia_FIC_PIC_compare.RData")
```

## Plot
```{r plot llk est by true parms, eval = F}
library(dplyr)
library(ggplot2)
rslt <- as.data.frame(rslt)
rslt$method[rslt$method == 1] <- "FIC"
rslt$method[rslt$method == 2] <- "PIC"
rslt$method[rslt$method == 3] <- "Vecchia"
rslt$method[rslt$method == 4] <- "SVecchia"
rslt$method[rslt$method == 5] <- "FITC"
rslt$method = factor(rslt$method, 
                     levels = c("FIC", "FITC", "PIC", "Vecchia", "SVecchia"))
mydf <- rslt %>% group_by(m, method) %>% 
  summarise(llk = mean(llk))

ggplot(data = mydf, aes(x = m, y = llk, group = method, col = method, 
                        label = method)) +
  geom_line() +
  xlab("m") +
  ylab("Avg. KL Divergence") +
  scale_color_brewer(palette = "Dark2") +
  # scale_y_continuous(trans = pseudo_log_trans(sigma = 1e-1),
  #                    breaks = unique(c(r, relInit))) +
  theme(legend.title = element_blank(), text = element_text(size = 20),
        legend.position="right")
ggsave("Vecchia_FIC_PIC_compare.pdf", width = 10, height = 5)
```


















