---
title: "Bridge GP Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Method 
Use mini-batching to fit models under different `lambda`. 

## Experiment
Clean the environment
```{r clean environment}
rm(list = ls())
```
Need to install a modified version of the GpGp R package.
```{r install the moderated GpGp Pkg}
devtools::install_github("https://github.com/SamCao1991/GpGp.git")
library(GpGp)
library(mvtnorm)
library(lhs)
library(FNN)
source("CQCD.R")
set.seed(123)
```
Simulate GP.
```{r simulate GP, eval = T}
n <- 1e4
d <- 1e3
m <- 100
k <- 3
batchSz <- 1e3
rSq <- c(10, 5, 2, 1, 0.5, rep(0, d - 5))^2
sigmasq <- 1.0 # variance
tausq <- 0.05^2 # nugget
locs <- randomLHS(n, d)
locs <- locs * outer(rep(sqrt(n), n), 
                     1 / sqrt(colSums(locs^2)))
covM <- matern25_scaledim_sqrelevance(c(sigmasq, rSq, tausq), locs)
cholM <- t(chol(covM))
y <- as.vector(cholM %*% rnorm(n))
X <- matrix(1, n, 1)
rm(covM, cholM)
```
Function for computing the approximate LOO-CV score, assuming the squared relevance parameterized Matern25 kernel.
```{r LOO-CV score, eval = T}
CV_score <- function(theta)
{
  idxLocsRel <- which(theta[2 : (d + 1)] > 0)
  locsRel <- locs[, idxLocsRel, drop = F]
  thetaRel <- theta[c(1, idxLocsRel + 1, d + 2)]
  dRel <- length(idxLocsRel)
  locsRelScal <- locsRel %*% 
    diag(sqrt(thetaRel[2 : (dRel + 1)]), dRel, dRel)
  NNarray <- get.knnx(locsRelScal, locsRelScal, m + 1)$nn.index
  - vecchia_profbeta_loglik(thetaRel, "matern25_scaledim_sqrelevance",
                            y, X, locsRel, NNarray)$loglik
}
```
Square-root L1 penalty functions.
```{r logistic penalty, eval = T}
gamma <- 1/4
penfun <- function(theta){
  lambda * sum(theta[-c(1, length(theta))]^(gamma))
}
dpenfun <- function(theta){
  r <- theta[-c(1, length(theta))]
  rpen <- lambda * r^(gamma - 1) * gamma
  rpen[r < 1e-10] <- lambda * (1e-10)^(gamma - 1) * gamma
  c(0, rpen, 0)
}
ddpenfun <- function(theta){
  diag(rep(0, length(theta)))
}
```
Function for model fitting given lambda.
```{r fit model for lambda, eval = T}
fit_order_select_SR <- function(theta0, idx0, gradObj0)
{
  theta <- theta0
  idx <- idx0
  gradObj <- gradObj0
  while(T)
  {
    odrDec <- order(gradObj$grad[2 : (d + 1)], decreasing = T)
    if(length(idx) == 0)
      idxNew <- 1 + odrDec[1 : k]
    else
      idxNew <- c(idx, 1 + setdiff(odrDec, idx - 1)[1 : k])
    idxSel <- setdiff(idxNew, idx)
    cat("Selected var:", idxSel, "\n")
    # opt with SR parameters
    thetaRel <- theta[c(1, idxNew, 2 + d)] 
    thetaRel[(length(idx) + 2) : (length(idx) + k + 1)] <- 1
    # minibatch
    idxBat <- sample(1 : n, batchSz, F)
    # maximin order and NNarray
    locsRel <- locs[idxBat, idxNew - 1, drop = F]
    locsRelScal <- locsRel %*% 
      diag(sqrt(thetaRel[2 : (1 + length(idxNew))]), 
           length(idxNew), length(idxNew))
    odr <- GpGp::order_maxmin(locsRelScal)
    yOdr <- y[idxBat[odr]]
    locsRelScalOdr <- locsRelScal[odr, , drop = F]
    locsOdr <- locs[idxBat[odr], , drop = F]
    XOdr <- X[idxBat[odr], , drop = F]
    NNarray <- GpGp::find_ordered_nn(locsRelScalOdr, m = m)
    # extract relevant col in locsOdr and lb
    locsOdrRel <- locsOdr[, idxNew - 1, drop = F] 
    lbRel <- lb[c(1, idxNew, 2 + d)] 
    objfun <- function(theta){
      likObj <- GpGp::vecchia_profbeta_loglik(theta, 
                                              "matern25_scaledim_sqrelevance",
                                              yOdr, XOdr, locsOdrRel, 
                                              NNarray)
      likObj$loglik <- likObj$loglik - penfun(theta)
      likObj
    }
    objfun_gdfm <- function(theta){
      likObj <- GpGp::vecchia_profbeta_loglik_grad_info(theta, 
                                              "matern25_scaledim_sqrelevance",
                                              yOdr, XOdr, locsOdrRel, 
                                              NNarray)
      likObj$loglik <- likObj$loglik - penfun(theta)
      likObj$grad <- likObj$grad - dpenfun(theta)
      likObj$info <- likObj$info + ddpenfun(theta)
      likObj
    }
    optObj <- CQCD(objfun, objfun_gdfm, thetaRel, maxIterOut = 50, 
                   maxIterIn = 40, lb = lbRel, arg_check = arg_check_SR)
    # Take out zero relevance
    thetaRel <- optObj$covparms
    idxLocZero <- which(thetaRel[2 : (1 + length(idxNew))] == 0)
    if(length(idxLocZero) > 0)
    {
      idxDesel <- idxNew[idxLocZero]
      cat("Predictor", idxDesel, "are zerod out \n")
      idxNew <- idxNew[-idxLocZero]
      thetaRel <- thetaRel[-(idxLocZero + 1)]
    }
    theta <- rep(0, length(theta))
    theta[c(1, idxNew, length(theta))] <- thetaRel
    idx <- idxNew
    # compute grad with thetaNew
    gradObj <- 
      GpGp::vecchia_profbeta_loglik_grad_info(theta, 
                                              "matern25_scaledim_sqrelevance",
                                              yOdr, XOdr, locsOdr,
                                              NNarray)
    # break condition
    if((length(idxLocZero) > 0 && any(idxSel %in% idxDesel)) || 
       length(idx) > 20)
      break
  }
  cat("lambda =", lambda, "idx =", idx, "\n\n")
  return(list(lambda = lambda, theta = theta, idx = idx, locsOdr = locsOdr, 
              yOdr = yOdr, XOdr = XOdr, NNarray = NNarray, gradObj = gradObj))
}
```

Storage variabels for plotting later.
```{r lambda grid, eval = T}
# define result collecting vars
idxSet <- list()
thetaSet <- list()
scoreSet <- list()
lambdaSet <- list()
loglikSet <- list()
```
Initialization.
```{r nitialization, eval = T}
sigmasqInit <- 0.25
tausqInit <- 0.01^2
theta <- c(sigmasqInit, rep(1e-8, d), tausqInit)
lb <- c(0.01^2, rep(0, d), 0.01^2) 
arg_check_SR <- function(x) {sum(sqrt(x[-c(1, length(x))])) > 1e-4}
arg_check_R <- function(x) {sum(x[-c(1, length(x))]) > 1e-4}
idx <- c()
XOdr <- X
yOdr <- y
locsOdr <- locs
NNarray <- find_ordered_nn(locs, m = m)
gradObj <- vecchia_profbeta_loglik_grad_info(theta, 
                                            "matern25_scaledim_sqrelevance",
                                            yOdr, XOdr, locsOdr, NNarray)
```
Loop over lambda.
```{r loop over lambda, eval = T}
lambdaVec <- rev(c(0, kronecker(10^(-1 : max(round(log10(batchSz)) - 2, 1)), 
                                c(1, 2, 5))))
niter <- length(lambdaVec)
for(i in 1 : niter)
{
  cat("\n====================================\n")
  lambda = lambdaVec[i]
  cat("i =", i, "lambda =", lambda, "\n")
  # fit model with penalty
  optObj <- fit_order_select_SR(theta, idx, gradObj)
  theta <- optObj$theta
  idx <- optObj$idx
  gradObj <- optObj$gradObj
  # Store results
  idxSet[[i]] <- optObj$idx
  thetaSet[[i]] <- optObj$theta
  scoreSet[[i]] <- CV_score(optObj$theta)
  lambdaSet[[i]] <- lambda
  loglikSet[[i]] <- optObj$gradObj$loglik
}
```

```{r save the result, eval = T}
save(list = c("d", "idxSet", "thetaSet", "scoreSet", "lambdaSet", "loglikSet",
              "lambdaVec", "n"), 
     file = paste0("bridge_minibat_", n, "_", d, ".RData"))
```

## Plot
Load libraries for plotting.
```{r lib for plotting, eval = F}
library(ggplot2)
library(RColorBrewer)
library(scales)
library(reshape)
```
Build dataframe for plotting theta.
```{r df theta, eval = F}
d <- length(thetaSet[[1]]) - 2
dfTheta <- matrix(NA, length(idxSet), d + 1)
colnames(dfTheta) <- c("lambda", paste0("var", 1 : d))
for(i in 1 : length(idxSet))
{
  dfTheta[i, 1] <- lambdaSet[[i]]
  dfTheta[i, 2 : (d + 1)] <- thetaSet[[i]][2 : (d + 1)]
}
dfTheta <- as.data.frame(dfTheta)
dfTheta <- melt(dfTheta, id = c(1))
colnames(dfTheta) <- c("lambda", "varID", "value")

dfThetaTrue <- data.frame("varID" = paste0("var", 1 : d), 
                          "trueVal" = c(100, 25, 4, 1, 0.25, rep(0, d - 5)))
```
Plot the change of `theta` v.s. `lambda`.
```{r plot theta-lambda, eval = F}
ggplot(data = dfTheta, aes(x = lambda, y = value, col = varID)) +
  geom_line() +
  geom_hline(mapping = aes(col = varID, yintercept = trueVal), 
             data = dfThetaTrue,
             size = 0.5, lty = "dashed") + 
  geom_hline(yintercept = 0, size = 0.5) + 
  scale_y_continuous(trans = pseudo_log_trans(sigma = 0.01),
                     breaks = unique(dfThetaTrue$trueVal)) +
  scale_x_continuous(trans = pseudo_log_trans(sigma = 1),
                     breaks = lambdaVec) +
  scale_color_manual(breaks = unique(dfThetaTrue$varID),
                     values = c(rainbow(5), rep("grey", d - 5))) +
  theme(legend.position = "none")
ggsave(paste0("theta_minibat_", n, "_", d, ".pdf"), width = 7, height = 5)
```

Plotting score.
```{r plot score, eval = F}
dfScore <- matrix(NA, length(idxSet), 2)
colnames(dfScore) <- c("lambda", "score")
for(i in 1 : length(idxSet))
{
  dfScore[i, 1] <- lambdaSet[[i]]
  dfScore[i, 2] <- scoreSet[[i]]
}
dfScore <- as.data.frame(dfScore)
dfScore$score <- dfScore$score - min(dfScore$score)

ggplot(data = dfScore, aes(x = lambda, y = score)) + 
  geom_line() +
  scale_x_continuous(trans = pseudo_log_trans(sigma = 0.1),
                     breaks = lambdaVec) +
  scale_y_continuous(trans = pseudo_log_trans(sigma = 0.1),
                     breaks = lambdaVec) +
  theme(legend.position = "none")
ggsave(paste0("score_minibat_", n, "_", d, ".pdf"), width = 7, height = 5)
```







