---
title: "Vecchia_FIC_PIC_compare"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r preparation}
rm(list = ls())
set.seed(123)
library(GpGp)
library(mvtnorm)
library(dplyr)
library(ggplot2)

n <- 5000
d <- 10
mVec <- seq(from = 4, to = 52, by = 4)
nrep <- 10
rTrue <- 20
rFalse <- 0.2
r <- c(rep(rTrue, 2), rep(rFalse, d - 2))
sigmaSq <- 1
tauSq <- 0
kernel_func <- matern25_scaledim_relevance
kernelName <- "matern25_scaledim_relevance"
```

## Function for simulating replicates
Function for simulating predictors and GP. Return a list consisting of locations `locs` and observations `y`.
```{r func for simulating predictors and GP}
simulate_func <- function(nExtra)
{
  nTtl <- n + nExtra
  locs <- lhs::randomLHS(nTtl, d)
  # locs <- locs * outer(rep(sqrt(nTtl), nTtl), 1 / sqrt(colSums(locs^2)))
  covM <- kernel_func(c(sigmaSq, r, tauSq), locs)
  cholM <- t(chol(covM))
  y <- as.vector(cholM %*% rnorm(nTtl))
  list(locs = locs, y = y)
}
```

## FIC and PIC
Function for selecting the inducing points (pseudo-inputs). Return the indices of inducing points.
```{r func for selecting inducing pts}
select_induce <- function(locs, m)
{
  order_maxmin(locs)[1 : m]
}
```
Function for partitioning the observations based on the `kmeans` function.
```{r func for partitioning}
partition_func <- function(locs, k)
{
  clusterID <- kmeans(locs, centers = k)$cluster
  partition <- list()
  for(i in 1 : k)
    partition[[i]] <- which(clusterID == i)
  partition
}
```
Function for computing the low-rank covariance matrices, returning the sum of the low-rank covariance matrix `Q` and the diagonal nugget.
```{r func for lr covM}
lr_covM <- function(locs, centers)
{
  covMTTl <- kernel_func(c(sigmaSq, r, tauSq), rbind(locs, centers))
  m1 <- nrow(locs)
  m2 <- nrow(centers)
  M <- covMTTl[(m1 + 1) : (m1 + m2), (m1 + 1) : (m1 + m2)]
  KXM <- covMTTl[1 : m1, (m1 + 1) : (m1 + m2)]
  Q <- KXM %*% solve(M) %*% t(KXM)
  diag(Q) <- diag(covMTTl)[1 : m1]
  Q
}
```
Function for the FIC log-likelihood. `locs` and `centers` should not overlap.
```{r func FIC log-likelihood}
loglk_FIC <- function(y, locs, centers)
{
  Q <- lr_covM(locs, centers)
  dmvnorm(y, sigma = Q, log = T)
}
```
Function for the PIC log-likelihood. `locs` and `centers` should not overlap.
```{r func PIC log-likelihood}
loglk_PIC <- function(y, locs, centers)
{
  partition <- partition_func(locs,
                              round(nrow(locs) / nrow(centers)))
  Q <- lr_covM(locs, centers)
  for(i in 1 : length(partition))
  {
    setI <- partition[[i]]
    if(length(setI) > 0)
    {
      Q[setI, setI] <- kernel_func(c(sigmaSq, r, tauSq), 
                                   locs[setI, , drop = F])
    }
  }
  dmvnorm(y, sigma = Q, log = T)
}
```

## Vecchia and Scaled Vecchia
Function for the Vecchia log-likelihood.
```{r func Vecchia log-likelihood}
loglk_Vecchia <- function(y, locs, m)
{
  odr <- order_maxmin(locs)
  locsOdr <- locs[odr, ]
  NNarray <- find_ordered_nn(locsOdr, m)
  yOdr <- y[odr]
  vecchia_meanzero_loglik(c(sigmaSq, r, tauSq), kernelName,
                          yOdr, locsOdr, NNarray)[[1]]
}
```
Function for the scaled Vecchia log-likelihood
```{r func scaled Vecchia log-likelihood}
loglk_SVecchia <- function(y, locs, m)
{
  odr <- order_maxmin(t(t(locs) * r))
  locsOdr <- locs[odr, ]
  NNarray <- find_ordered_nn(t(t(locsOdr) * r), m)
  yOdr <- y[odr]
  vecchia_meanzero_loglik(c(sigmaSq, r, tauSq), kernelName, 
                          yOdr, locsOdr, NNarray)[[1]]
}
```

## Simulation
```{r simulation}
rslt <- matrix(NA, 4 * length(mVec) * nrep, 4)
colnames(rslt) <- c("iter", "m", "method", "llk")
rslt[, 3] <- c(1, 2, 3, 4) # method ID
rslt[, 2] <- kronecker(mVec, rep(1, 4)) # m
rslt[, 1] <- kronecker(c(1 : nrep), rep(1, 4 * length(mVec))) # iter ID
offset <- 0
for(i in 1 : nrep)
{
  mMax <- max(mVec)
  simObj <- simulate_func(mMax)
  idxInduce <- select_induce(simObj$locs, mMax)
  locs <- simObj$locs[- idxInduce, ]
  # true density
  trueDen <- dmvnorm(rep(0, n),
                     sigma = kernel_func(c(sigmaSq, r, tauSq), locs),
                     log = T)
  for(m in mVec)
  {
    locsInduceFIC <- simObj$locs[idxInduce[1 : m], ]
    locsInducePIC <- simObj$locs[idxInduce[1 : (m / 2)], ]
    
    rslt[offset + 1, 4] <- trueDen - loglk_FIC(rep(0, n), locs, locsInduceFIC)
    rslt[offset + 2, 4] <- trueDen - loglk_PIC(rep(0, n), locs, locsInducePIC)
    rslt[offset + 3, 4] <- trueDen - loglk_Vecchia(rep(0, n), locs, m)
    rslt[offset + 4, 4] <- trueDen - loglk_SVecchia(rep(0, n), locs, m)
    
    offset <- offset + 4
  }
}
save(rslt, file = "Vecchia_FIC_PIC_compare.RData")
```

## Plot
```{r plot llk est by true parms}
rslt <- as.data.frame(rslt)
rslt$method[rslt$method == 1] <- "FIC"
rslt$method[rslt$method == 2] <- "PIC"
rslt$method[rslt$method == 3] <- "Vecchia"
rslt$method[rslt$method == 4] <- "SVecchia"
rslt$method = factor(rslt$method, 
                     levels = c("FIC", "PIC", "Vecchia", "SVecchia"))
mydf <- rslt %>% group_by(m, method) %>% 
  summarise(llk = mean(llk))

ggplot(data = mydf, aes(x = m, y = llk, group = method, col = method, 
                        label = method)) +
  geom_line() +
  xlab("m") +
  ylab("Avg. KL Divergence") +
  scale_color_brewer(palette = "Dark2") +
  # scale_y_continuous(trans = pseudo_log_trans(sigma = 1e-1),
  #                    breaks = unique(c(r, relInit))) +
  theme(legend.title = element_blank(), legend.position="right")
ggsave("Vecchia_FIC_PIC_compare.pdf", width = 6, height = 3.375)
```


















